### 2020-05-16 leadership_across_cultures_contexts.Rmd
Adding text to SI.

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: Leader2.bib
csl: apa-6th-edition.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
library(readr)
library(knitr)
library(pander)
library(patchwork)
library(gt)

# Need there here so they are identified as dependencies of this report

loadd(all_data)
loadd(all_data2)
loadd(all_study_vars)
loadd(coercive_plot)
loadd(df_func)
loadd(df_qual)
loadd(df_shaman)
loadd(features)
loadd(feature_models)
loadd(feature_models_aic)
loadd(features_support_cult)
loadd(follower_bc_ratio)
loadd(follower_benefits_txt)
loadd(follower_costs_txt)
loadd(follower_benefits_cult)
loadd(follower_costs_cult)
loadd(functions_support_cult)
loadd(functions_support_txt)
loadd(highstatus_plot)
loadd(leader_bc_ratio)
loadd(leader_benefits_txt)
loadd(leader_benefits_cult)
loadd(leader_costs_txt)
loadd(leader_costs_cult)
loadd(m_lpca_func)
loadd(m_lpca_qual)
loadd(m_pvclust_func)
loadd(m_pvclust_qual)
loadd(multi_aic)
loadd(multi_allvars)
# loadd(p_heatmap_groups)
# loadd(p_heatmap_region)
# loadd(p_heatmap_sex)
# loadd(p_heatmap_subsis)
loadd(p_heatmap_feature_subsis)
loadd(p_heatmap_feature_region)
loadd(p_heatmap_feature_sex)
loadd(p_heatmap_feature_groups)
loadd(plot_costs_benefits)
loadd(plot_elastic_shamanism)
loadd(plot_elastic_status)
loadd(plot_elastic_coercive)
loadd(plot_features_support)
loadd(plot_func_qual)
loadd(plot_pubdate)
loadd(plot_shamanism_text)
loadd(qualities_support_cult)
loadd(qualities_support_txt)

source("R/cheapobjects.R", local = T)
```

<!-- A thematic note: One of our primary goals is to explore how an evolutionary framework can illuminate when we expect to find cultural, behavioral, physiological, etc. variation and then explain (some of) that variation. Part of going "Beyond WEIRD" is getting to theories that can account for population-level variation and predict invariant universals (or, at least, less variable phenotypes). We want to push things beyond the 'butterfly collecting' of simply documenting population-level variation. -->

# Abstract

Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their complex organizations. Many researchers have turned to evolutionary theory to better understand this diversity. Evolutionary theories of leadership, in turn, draw on ethnographic cases of societies thought to more closely resemble the smaller-scale, face-to-face communities in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

Building on our previous cross-cultural findings we systematically evaluate dimensions of leadership across a broad range of human cultural diversity. Over 1200 ethnographic texts from 59 mostly non-WEIRD populations in Human Relations Area Files (HRAF) were coded on leader qualities and functions. Because adaptations for leadership and followership could only have evolved if the fitness benefits exceeded the costs, we coded the costs and benefits for both leaders and followers across diverse contexts. We discovered evidence for both cultural universals in leadership, as well as important variation. Leaders in the ethnographic record were most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial. The primary benefits leaders offered to followers were resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources. The primary benefits leaders received from followers included material resources, increased social status, and social services. The most frequent costs of leadership included loss of social status, resources or social services and increased risk of harm or conflict. For followers, the most frequent costs included loss of resources, costs in providing social services, and increased risk of harm or conflict.

Shamans emerged as an important category of leaders who do not clearly conform to influential models that emphasize two leader strategies: using knowledge and expertise to provide benefits to followers vs. using physical formidability to impose costs. Instead, shamans and other leaders with supernatural abilities use their knowledge to both provide benefits and impose costs on others. We therefore propose a modified scheme in which leaders deploy their cognitive, social, and somatic capital to provide benefits and/or impose costs on others.

Leadership in WEIRD settings often differs substantially from non-WEIRD settings, where influence crosses domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face. The diversity within the non-WEIRD, however, should not be overlooked. Leadership theories, including evolutionary and managerial approaches have not previously had broad systematic, cross-cultural data on leadership, management, and social organization across diverse populations. A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.

# Introduction

>It's good to be king, if just for a while. - Tom Petty

<!-- INTRO TO MOTIVATE PAPER -->

Research on leadership within anthropology extends to the birth of the discipline. Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917]. Anthropologists have documented the emergence of leaders among every ethnographically observed society [@brown_human_1991; @lewis_leaders_1974], and in diverse contexts, such as kin groups [@dussart2000politics], ritual [@singh_cultural_2017], economic groups [@macfarlan_competitive_2012], conflicts between groups [@glowacki_evolutionary_2017], and in nonindustrial political groups [@cohen_comparative_1967]. For review, see @garfield_review_2019.

Despite decades of research and theorizing on leadership in numerous social science disciplines, to our knowledge there is no comprehensive or systematic investigation of leader functions and qualities, and the costs and benefits of leadership and followership, across a representative sample of nonindustrial societies. Moreover, outside of anthropology, most research and theorizing on leadership is based on data from Western or Westernized, postindustrial societies [@von_rueden_leadership_2015].

<!-- WEIRD PEOPLE PROBLEMS -->
There is now broad consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*]. Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

Anthropology, on the other hand, has been criticized for exoticizing and essentializing diverse "Others" as part of its colonial history [@said1979; @abu2008; @reyes-garcia_small-scale_2017]. @abu2008 [p. 470], for instance, notes that

>Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor -- race -- even though in its twentieth-century form it has some important political advantages.

<!-- culture is important to anthropology because the anthropological distinction between self and other rests on it. Culture is the essential tool for making other. As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference, anthropology also helps construct, produce, and maintain it. Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages. Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences. -->

There is potentially as much diversity within nonindustrial, non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991]. @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies. It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.

<!-- Extending the study of leadership to include evolutionary theory and data from diverse, small-scale and nonindustrial societies (i.e., non-WEIRD) can facilitate more generalizable theories and offer applications to shape organizational structure across populations (i.e., WEIRD settings) in ways which are more consistent and less mismatched with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010]. Such approaches can also incorporate the anthropological insight that diversity of cultural, social, and ecological contexts are strong drivers of human behavior and group dynamics [@wiessner_risk_1982; @henrich_big_2015]. -->

<!-- as well as the costs and benefits to individuals involved. Cultural norms, capacities of competing leaders, inter-group dynamics, and follower preferences can also impact leader agendas. The interactions between leader strategies, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories  and cultural and biological evolutionary processes [].   -->

<!-- EVOLUTIONARY THEORY APPROACH -->
<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex. Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019]. This integration as been informative [see @van_vugt_genes_2020]. -->

Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be. One outstanding question for evolutionary leadership scholars is the degree to which theoretically important dimensions of leadership vary between cultures and across social contexts. Leadership studies, including evolutionary approaches, have focused on community or political leaders, often overlooking the nature of leadership and followership across social contexts and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities. A second outstanding question is what benefits *and* costs do both leaders and followers incur. While the benefits leaders accrue for their services have received substantial attention -- fitness benefits in particular -- the costs of leadership for both leaders and followers have been relatively under-investigated [for discussions on costs of leadership see @cheng_dominance_2019; @price_evolution_2014; @glowacki_leadership_2015; @von_rueden_leadership_2014].

<!-- Leadership roles are also often associated with socially sanctioned rights and responsibilities. Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face. Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them. Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families. Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].  -->

<!-- CLOSING -->
We first briefly review the qualities of leaders, the functions they serve, and the costs and benefits of leadership across nonindustrial, non-WEIRD human societies. We then outline the aims and methods of the current cross-cultural study -- which focuses on exploring patterns among non-WEIRD populations -- before presenting results and discussion.

## The functions of leadership across cultures and contexts
<!-- INTRO -->
@roscoe_social_2009, drawing on the ethnography of highland New Guinea, proposes that leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling military formidability to rival groups. Among many ethnographically described small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when group demands require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996]. Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Representing the group and managing internal and external relationships and conflicts are expected to be the most salient functions of smaller-scale community leaders.

<!-- GETS INTO QUALITIES BELOW -->
<!-- Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].  -->

<!-- MANAGEMENT FUNCTION IN MORE CULTUARLL COMPLEX -->
Increased sedentarization and defensibility of resources is often associated with social stratification, broader institutionalized managerial roles, and stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004]. Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014]. Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990]. 
<!-- In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.  -->
Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017]. Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military operations [@lopez_evolutionary_2017].

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016]. 
<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.  -->
Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics. We include detiled review of this literature on WEIRD populations in the SI.

## The qualities of leaders across cultures and contexts

<!-- UNIVERSAL QUALITIES OF COMMUNITY LEADERS -->
There has been increasing interdisciplinary focus on the closely related constructs of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014]. Whereas leaders have disproportionate group influence, status is often conceptualized as a relative indicator of the social value of the individual by the group and of their relative access to contested resources [e.g., @blader_whats_2014; @von_rueden_roots_2014]. The relationship between leadership and status is often expected to be colinear [e.g., @von_rueden_leadership_2015; @van_vugt_genes_2020; @cheng_dominance_2019; @cheng_psychology_2014]. The causal relationship between the two, however, is likely to be complex, context-specific, and driven by underlying correlates. Community leaders are frequently conceptualized as high status males who embody the qualities most respected by the group and necessary for successful leadership in the contexts in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].

<!-- MALE BIAS -->
Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978]. @low_sex_1992 demonstrated in a systematic cross-cultural sample that political leaders of both sexes were present in only seven of 93 societies and in only two societies were female political leaders identified as equally powerful as their male counterparts. Even among these rare cases, powerful female political leaders are still infrequent within the group and women were more likely to hold influential leadership positions within kin groups [@low_sex_1992]. @yanca_female_2004 found that sex-differences in political influence was more likely driven by ecological and social factors, such as social networks and access to same-sex allies, rather than sex (or gender identity) per se. Two recent studies relying on ethnographic fieldwork comparing leaders and non-leaders in small-scale societies, both revealed sex is not a strong predictor of community influence or elected leadership, after accounting for economic specializations and a broad range of demographic, phenotypic, and social traits [@garfield_investigating_2019; @von_rueden_sex_2018].

<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
Across cultures elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].

<!-- NON WEIRD QUALITIES -->
Among small-scale, autonomous, relatively egalitarian societies community leaders tend to be older, respected males, with large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012]. Physically, leaders across multiple contexts tend to be taller and more physically formidable. Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014]. Charisma and oratory skills are common properties of leaders, but thought to be especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].

<!-- ASCRIBED THINGS -->
In the context of institutionalized social stratification, such as among chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998]. Managerial elites maintain influence over community decisions and are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010]. In such contexts community leaders often lack strong coercive authority, but maintain influence through ideological systems and ritual or religious statuses inaccessible to non-elites [@stanish_evolution_2010]. Within state-level societies including nonindustrial kingdoms leadership positions are proscribed and institutionalized. Ruling classes, parties, or political bodies monopolize social political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].

<!-- INTELLIGENCE -->
Many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], including across diverse societies [@garfield_evolutionary_2019; @garfield_investigating_2019]. These results suggests high intelligence and specialized knowledge are potentially universal traits of political and community leaders.

<!--  CONCLUSION PARAGRAPH/STATEMENT. -->
Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders. Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be broadly representative qualities of political and community leaders.

<!-- We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities. Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use. In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not. In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.    -->

## The benefits and costs of leadership across cultures and contexts

<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
The ability to influence group processes can provide enormous benefits to leaders, even if 'just for a while.' Leaders in any context can promote policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019]. 
<!-- BENEFITS: HIGH SOCIAL STATUS -->
Leadership positions, including formal and informal roles, are commonly associated with high social status and numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014]. Leaders and high-status individuals are more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004]. Across many WEIRD organizations leadership positions are associated with increased financial compensation [@tsui_income_2017]. High social status, material wealth, intelligence, and leadership status are also generally viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].

<!-- BENFITS: MATING -->
Across diverse populations and organizations, community and political leaders, tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success. In a phylogenetic multilevel meta-analysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.

<!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.

<!-- COSTS: OPPURTUNITY COSTS -->
Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers, on average. However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership. Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources. 
<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010]. The solutions leaders propose are unlikely to be equally satisfactory for all followers and leaders can become the targets of animosity. In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017]. Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts. Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015]. With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude. In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.

# Study aims

Evolutionary theoretical models and studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. We seek to provide a systematic and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample of nonindustrial, non-WEIRD populations. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how these dimensions of leadership vary within populations across social contexts and across populations with different subsistence strategies, as well as by leader sex and continental region.

In sum, we sought to empirically and systematically accomplish the following four goals: 1) identify variation in the *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular qualities or functions tend to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010]. 3) Identify variation in the *costs and benefits* of leadership for both leaders and followers. And 4) determine the extent to which the qualities and functions of leaders, and their costs and benefits, varied across social contexts within populations, varied across populations with different subsistence strategies, and varied by leader sex or continental region.

Our approach was exploratory: we did not have *a priori* hypotheses to test.

<!-- Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization. -->

# Methods

## Ethnographic sample and coding

We estimated the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 tested three theoretical models against the ethnographic record. The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. The eHRAF is a continually growing subset of the larger HRAF.

Whereas @garfield_evolutionary_2019 documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The PSF aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. It comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text.

The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.

```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
```

In contrast to the qualities and functions variables, coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology. For both followers and leaders, we coded if there was evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories. See Table \@ref(tab:costbenefitvars). There are multiple instances of conceptual overlap between many of our researcher-coded variables, in particular between various qualities and related costs or benefits. The operationalization of costs and benefits generally capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description. We return to this point in the limitations section.

```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
```

ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS coded all text records for group structure.

ZG and KS then discussed every instance of disagreement between their respective coding on all variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:

>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief. The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e. those in authority over each community, consisting of from 20 to 50 individuals, maicla. The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited. It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers. It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating. It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved. The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.

This text record was coded has having evidence for the functions *Resolve conflict*, *Organize cooperation*, *Movement/migration*, *Provide subsistence*, and *Misc. social functions*. For qualities, there was evidence for *High status*. Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization.

Each culture was coded for subsistence strategy. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.

## Statistical analysis

Our data comprised one row for each of our text records. For each of our variables, each row was coded as 1 if there was evidence in the text record for that variable, and 0 otherwise. Because absence of evidence is not evidence of absence, our data only inform the extent to which there is evidence for each of the operationalized leader traits.

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record. We computed the proportion of text records supporting each variable, and their 95% confidence intervals, using logistic generalized linear mixed effects models with random intercepts for document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework). For each variable, we then computed the proportion of cultures with at least one text record with evidence for that variable, with confidence intervals computed using a cluster bootstrap.

Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.

Our third goal was to assess the extent to which each of the researcher-coded dimensions of leadership varied by four theoretically important factors for which we had data for all rows: group context, subsistence strategy, continental region, and leader sex.

Our fourth goal was to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].

NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$. NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].

Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster. We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

Our fifth goal was to assess variation in our lower-dimensional "features" by group context, subsistence strategy, continental region, and leader sex.

Our sixth goal was to evaluate the features in light of influential evolutionary theories of leadership. Drawing on the empirically derived features as a guide, we aimed to construct "elements" of leadership that corresponded to elements of influential theories.

<!-- To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(df_qual)` text records for the leader quality variables and `r nrow(df_func)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors. -->

<!-- To identify natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster analyses using the pvclust package [@suzuki_pvclust:_2015]. These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary. -->

To supplement results from the aforementioned methods, we also performed textual analyses and explored variation measures of biases stemming from the ethnographic record that may have impacted results.

<!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups. We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts. -->

The analyses computing levels of evidence for each variable used the entire set of text records. Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any variable (i.e., all zeros), and variables with very little evidence (almost all zeros). All analyses were conducted with `r R.version.string`.

# Results

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Figure S\@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

```{r corpus}
cor_pages_tr <- cor.test(~number_leader_records + ehraf_pages, leader_cult)
```

<!-- NEED TO DISCUSS 1187 and split records.  -->

## Distribution of text records by subsistence strategy and group type

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplementary Figure \@ref(fig:regionsubsistence)).

Subsistence strategy and group structure were also confounded. Hunter-gatherer text records often referred to residential leadership (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and kin leadership (`r hg_kin_pct`%), but rarely referred to supracommunity leadership (`r hg_supra_pct`%). Agricultural text records, on the other hand, rarely referring to residential leadership (`r agri_residential_pct`%) but often referring to supracommunity leadership (`r agri_supra_pct`%). Across all subsistence types, the most frequent group type was political groups at the community level, with the exception of agriculturalists in which supracommunity political groups are the most frequent group type. Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers. See Figure \@ref(fig:grouptype).

```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group context. Areas of the bars are proportional to the number of text records in that category."}
ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
plot_group_subsis
```

## Leader sex

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders. The majority of text records which exclusivley described female leaders were from horticulturlaists (`r female_hort_pct`%). See Table S\@ref(tab:sexsubtable).

<!-- SHOULD WE PUT THE ENTIRE TABLE IN SI? -->

## Possible sources of bias

```{r tabauthorgender}
unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
tab_author_gender <- table(unique_authors$author_gender)
mm_coauthorSUM <- summary(mm_coauthor)
mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
```

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

The number of text records on leadership retrieved for a particular culture was weakly correlated with the total number of pages of ethnography in the eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). A loess regression suggested that the number of text records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure S\@ref(fig:eHRAFpages)).

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure S\@ref(fig:pubdate) and supplementary information for details.

We then tested if text records with a female author or co-author were more likely to discuss female leaders. Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership. Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

## Evidence for leader functions and qualities

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts. We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap). See Figure \@ref(fig:varsupport). For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).

```{r varsupport, fig.width=15, fig.height=15, fig.cap="Evidence for each coded variable of leader functions (top) and qualities (bottom). Red: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture). Purple: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
ggsave(file="Figures/variablesupport.pdf", plot_func_qual)
plot_func_qual
```

The most common functions of leaders, documented in over 70% of cultures, were resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs. Leaders were also commonly described as serving as a representatives in inter-group interactions, managing the distribution of resources, and providing context-specific leadership such as military command and in providing ritual services. In about half of the cultures sampled here leaders punished other group members and managed economic systems. Notable infrequent functions included controlling group movement and immigration.

The most common qualities of leaders were having high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures). Also, in about 60% of cultures sampled leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained coercive authority over followers. Notable infrequent qualities of leaders included physical formidability, having many social contacts, and being feared (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).

Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor). See Figures \@ref(fig:FunctionRE) and \@ref(fig:QualityRE). In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

## Benefits and costs of leadership and followership

```{r cultbc}
leader_benefit_material <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Material resources'], 3)

leader_benefit_status <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Increased social status'], 3)

leader_cost_material <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Loss of material resources'], 3)

leader_cost_status <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Reduced social status'], 3)

follower_benefit_material <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Material resources'], 3)

follower_benefit_social <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Social services'], 3)

follower_cost_material <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of material resources'], 3)

follower_cost_social <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of social services'], 3)
```

The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above. In these data leaders most commonly received benefits in the form of greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures). Mating benefits and social services benefits were also relatively frequent benefits of leadership. In general, the costs tracked the benefits. Leaders were often described as losing their high social status (documented in `r leader_cost_status`% of cultures) and greater access to resources (documented in `r leader_cost_material`% of cultures). See Figure \@ref(fig:varsupportcb).

Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in `r follower_benefit_material`% of cultures) but otherwise most frequently through receiving social services (documented in `r follower_benefit_social`% of cultures). Mating and fitness benefits were relatively rare for followers. Followers were most frequently described as incurring costs in the form of material resources lost (documented in `r follower_cost_material`% of cultures) and in providing social services (documented in `r follower_cost_social`% of cultures). See Figure \@ref(fig:varsupportcb).

```{r varsupportcb, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership. Blue: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture). Red: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
ggsave(file="Figures/variablecostsbenefits.pdf", plot_costs_benefits)
plot_costs_benefits
```

Leadership can evolve when the benefits exceed the costs, and the same for followership. Our data did not allow us to compare benefits and costs. Instead, we compared levels of evidence for benefits vs. costs for each leader benefit/cost dimension and each follower benefit/cost dimension. Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., *Mating*, *Food*), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture. We fit separate models for leaders and followers. These two models allowed us to estimate the relative evidence for, e.g., leader mating benefits vs. leader mating costs, and follower mating benefits vs. follower mating costs, etc. For leaders, the odds ratio that a text record would provide evidence for a benefit vs. a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$. Hence for both leaders and followers, there was more evidence of benefits than costs.

Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from harm vs. risk of harm. For leaders compared to followers, there were noticeably higher odds ratios for mating, inclusive fitness, material resources, and social services. See Figure \@ref(fig:benefitcostratio).

```{r benefitcostratio, fig.width=10, fig.cap="The odds ratios of finding evidence of a benefit vs. a cost in a text record for leaders and followers, by each dimension of benefits and costs. Odds ratios estimated using a mixed effects logistic regression model of evidence for (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture. Separate models were fit for leaders and followers. X-axis is on a log scale."}
ggsave(file='Figures/plot_bc_OR.pdf', plot_bc_OR, width = 9, height = 5)
plot_bc_OR
```

## Universality in leadership dimensions

To investigate variation in our `r length(all_study_vars)` dimensions of leadership, we fit a multiple logistic mixed effects regression model of each dimension as a function of four factors: group context, subsistence strategy, continental region, and leader sex, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts. Models of leadership dimensions whose AIC values that showed less than a two-point reduction relative to the intercept-only model were deemed to not vary by these factors, whereas those whose AIC values were less two or more were deemed to vary.

```{r universals}

commondim_n <- sum(df_compare2$Estimate > 0.6)
universals_n <- sum(df_compare2$AIC_diff >= -2 & df_compare2$Estimate > 0.6)
vary_n <- sum(df_compare2$AIC_diff < -2)

```

We first examined the `r commondim_n` leadership dimensions for which evidence was found in at least 60% of cultures. Of these, `r universals_n` did not meaningfully vary by the four factors. Because evidence for these dimensions was common across cultures, the `r universals_n` unvarying dimension are candidates for universal dimensions of human leadership. See Table \@ref(tab:leaderuniversals).

```{r leaderuniversals}

df_compare2 %>% 
  filter(AIC_diff >= -2 & Estimate > 0.6) %>% 
  mutate(
    Type = map_chr(var, ~ str_split(., '_')[[1]][[1]]),
    Type = str_replace(Type, '\\.', ' '),
    Type = str_to_title(Type),
    `Percent of cultures` = 100*Estimate
  ) %>% 
  select(Type, Variable, `Percent of cultures`) %>% 
  arrange(-`Percent of cultures`) %>% 
  kable(digits = 1, caption = "Candidate universal dimensions of leadership. These dimensions were common in the ethnographic record (appearing in at least 60% of cultures), and did not appear to meaningfully vary by group context, subsistence strategy, continental region, or leader sex.")

```

## Variation by group structure, subsistence strategy, continental region, and leader sex

We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)). To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC. For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others. Variation in group context, in particular, was associated with variation in almost half of the dimensions. See Figure \@ref(fig:heatmapdrop1)A.

```{r heatmapdrop1, fig.width=12, fig.cap="Leadership dimensions and features that varied by group context, subsistence strategy, continental region, and/or leader sex. A. Leadership dimensions. B. Leadership features. Colors represent the relative change in AIC value after dropping that factor from a multiple logistic mixed effects regression model. Yellow: including factor improved fit relative to other factors. Black: including factor did not improve fit relative to other factors. Delta AIC values were standardized by rows, and therefore can only be compared within rows."}

p <- heatmap_drop1 + guides(fill='none') + heatmap_features_drop1 + plot_annotation(tag_levels = 'A')
ggsave(filename = 'Figures/heatmap_drop1.pdf', p, width = 14, height = 5)

p
```

For each of our four factors, we then investigated which levels of the factor was associated with higher or lower levels of evidence for the leadership dimension. For instance, for subsistence strategy, there were three leadership dimensions in which it played an important role: *No coercive authority*, *Provide subsistence*, and *Distribute resources* (See Figure \@ref(fig:heatmapdrop1)). Hunter-gatherers had particularly high levels of evidence for *No coercive authority*, relative to the other subsistence strategies (and averaging over the levels of the other three factors), and Pastoralists had relatively low evidence for *Distribute resources* (see Figure S\@ref(fig:heatmapsubsis)). For Group context, see Figure S\@ref(fig:heatmapgroups); for Region, see Figure, S\@ref(fig:heatmapregion); and for Leader sex, see Figure S\@ref(fig:heatmapsex).

<!-- For subsistence strategy, hunter-gatherers had relatively more evidence for *Polygyny*, *Leader-Mating benefits*, *No coercive authority*, *Political appointments* and *Military command* than other subsistence strategies. Pastoralists had more evidence for *Ritual functions*, *Supernatural*, *Aggressiveness* and *High quality spouse*. Hunter-gatherers and pastoralists both had more evidence for *High status*, *Providing counsel/direction*, and *Providing subsistence*. The remaining subsistence strategies generally had relatively less evidence for all variables, except that the mixed strategy had relatively more evidence for *Organize cooperation*, *Distribute resources*, and *Prosocial investment*. See Figure S\@ref(fig:heatmapsubsis). -->

<!-- For continental region, there was little strong patterning, except that Africa generally had relatively less evidence for any retained variable except *Provide counsel/direction*, and South America had relatively more evidence for several variables, including *Age*, *Follower-Social services benefits*, and *High status*. See Figure S\@ref(fig:heatmapregion). -->

<!-- There was strong patterning of relative evidence for almost all variables by leader sex. Female leaders had relatively more evidence for *High quality spouse*, *Leader - Mating benefits*, *Polygyny*, *High status*, and *Age*. Male leaders had relatively more evidence for *Follower - Social  services benefits*, *Political appointments*, *Prosocial investment*, and *Leader-Social services costs*. Records pertaining to both sexes had relatively more evidence for *Organize cooperation*, *Providing subsistence*, *Ritual functions*, *Bravery*, *Experience/accomplished*, and few that overlapped with males, such as *Military command* and *Distributing resources*. See Figure S\@ref(fig:heatmapsex). -->

<!-- There was also strong patterning of relative evidence for almost all variables by group structure. There was relatively more evidence for military traits in military groups, relatively more evidence for ritual and supernatural traits in religious groups, relatively more evidence for organizing cooperation and providing subsistence in economic groups, and so forth. Kin groups had relatively more evidence for *Provide counsel/direction*, *Prosocial investment*, *Age*, and *High status*. Community and residential groups had relatively low evidence for most variables, except that there was relatively more evidence for *High quality spouses* among residential group leaders. See Figure S\@ref(fig:heatmapgroups). -->

## When do leaders have high status?

```{r highstatuscalc}
statusFracTxt <- qualities_support_txt$Estimate[qualities_support_txt$vars == 'qualities_HighStatus']
statusPercTxt <- signif(100*statusFracTxt, 3)

statusFracCult <- qualities_support_cult$Estimate[qualities_support_txt$vars == 'qualities_HighStatus']
statusPercCult <- signif(100*statusFracCult, 3)

```

*High status* was the most frequently identified leader quality in the sample, with `r statusPercTxt`% of text records and `r statusPercCult`% of cultures containing evidence that leaders had high social status. To explore factors that distinguished these records from those with no evidence for high status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occurred in each text record (see supplementary information for details). We then fit an elasticnet logistic regression model (with the lasso penalty, $\alpha = 1$) of *High status* (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words. Words that were strong positive predictors epitomized the semantic content of the text records which provided evidence of high status as a quality of leaders. These included *`r rev(highstatus_plot$data$Word)[1]`* and *`r rev(highstatus_plot$data$Word)[2]`*, which were the strongest predictive terms; *`r rev(highstatus_plot$data$Word)[3]`* and *`r rev(highstatus_plot$data$Word)[4]`* were also relatively strong predictors; and *`r rev(highstatus_plot$data$Word)[5]`* and *`r rev(highstatus_plot$data$Word)[6]`*, which were also positive predictors. Notable non-predictor terms included, *influence*, *authority*, *headman*, and *chief*. See Figure \@ref(fig:highstatusplot)A.

To determine which leader qualities and functions were associated with evidence for high status leaders, we fit an elasticnet logistic model of *High status* as a function of all quality and function variables, (with the lasso penalty, $\alpha = 1$). The non-zero coefficients indicated that *Punishment*, the only identified function, was a negative predictor, whereas
<!-- all other functions, including *Organize cooperation*, *Ritual functions*, and *Provide counsel/direction* were non-predictors, and  -->
*Wealthy*, *Strategic nepotism*, *Age*, *Experienced/accomplished*, and several other qualities, were positive predictors. See Figure \@ref(fig:highstatusplot)B.

```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*. Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

highstatus_plot <- highstatus_plot + scale_x_log10()
p <- 
  highstatus_plot + plot_elastic_status + 
  plot_annotation(
    title = 'Predictors of High status',
    tag_levels = 'A', 
    caption = 'Odds ratio', 
    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
  )

ggsave('Figures/highstatus_plot.pdf', p, width=10, height = 7)

p
```

We also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts). There were no clear differences. See Figure S\@ref(fig:groupstatus).

## When do leaders have coercive authority?

*Coercive authority* was documented in about 57% of cultures. Unlike other more frequent leader qualities, which were generally positive or prosocial traits, *Coercive authority* was the most frequent negative or dominance-style leader quality. We therefore chose to perform additional exploratory analyses with *Coercive authority* as the outcome in logistic elasticnet regression models similarly as described above for *High status* (see Figure \@ref(fig:elasticcoercive) caption for details).

```{r}
nocoerciveauthority <- signif(100*qualities_support_cult$Estimate[qualities_support_cult$Variable == 'No coercive authority'], 3)
```

Words that were strong predictors of leaders with coercive authority included, *power* and *authority* (unsurprisingly), as well as *territory*, *chief*, and *control*; the word *leadership* was the only strong negative predictor. See Figure \@ref(fig:elasticcoercive)A. Using all other quality and function variables as predictors identified *Punishment* as the strongest predictor of evidence for *Coercive authority*. The qualities of *Aggressiveness* and *Feared* were also positive predictors, as were the functions, *Political appointments* and *Control economics*. Furthermore, evidence for leaders lacking coercive authority (our *No coercive authority* variable) was documented in `r nocoerciveauthority`% of cultures (Figure \@ref(fig:varsupport)) and demonstrated variability by subsistence strategy with hunter-gatherers providing relatively higher levels of evidence for leaders lacking coercive authority. See Figure S\@ref(fig:heatmapdrop1).

```{r elasticcoercive, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Coercive authority*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Coercive authority* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Coercive authority*. Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

coercive_plot <- coercive_plot + scale_x_log10()
p <-
  coercive_plot + plot_elastic_coercive +
  plot_annotation(
    title = 'Predictors of Coercive authority',
    tag_levels = 'A', 
    caption = 'Odds ratio', 
    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
    )
ggsave('Figures/coercive_plot.pdf', p, width = 8, height = 4)
p
```

## Dimension reduction

```{r}
num_removed <- sum(rowSums(all_data[all_study_vars]) == 0)
```

Our qualities and functions data comprised sparse $`r nrow(all_data)` \times `r length(quality_vars)`$ and $`r nrow(all_data)` \times `r length(function_vars)`$ matrices of binary data, respectively. Most data reduction methods did not find strong or stable evidence of structure in the data. We therefore explored reduced matrices that removed uninformative rows with no evidence for any variable, and variables with only a few rows with evidence. Stable clusters emerged with $`r nrow(df_qual)` \times `r ncol(df_qual)`$ and $`r nrow(df_func)` \times `r ncol(df_func)`$ matrices for qualities and functions, respectively.

Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from other methods, see the supplementary information). Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $\gt95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

```{r cluster, fig.cap="Cluster analysis of (A) leader quality variables and (B) leader function variables. Distances were $1-cor$. Ward agglomeration method. AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to prosocial qualities and one that comprised other individual-level qualities which were not necessarily prosocial. Within the latter top-level cluster we identified two interpretable and strongly supported subclusters which are annotated with text (*Social, reproductive, and material success* and *Competencies*). Within the first top-level cluster, we also identified two subclusters, one of which contained measures related to group-level pressures (*Cultural conformity*) and one which contained the prosocial traits (*Prosocial competencies*); these subclusters were moderately supported.

The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (e.g., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (e.g., *Social functions*, *Distribute resources*, *Hospitality*). Within the management cluster, we identified two moderately supported subclusters; a *Mediation* subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an *Organization* subcluster (e.g., *Organize cooperation*, *Military command*).

We identify these seven named subclusters as "features" of leader qualities and functions in the ethnographic record. In naming these features, we gave extra weight to variables that were strongly supported in the sample (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes of each feature. These features correspond fairly well to major elements of evolutionary theories of leadership reviewed in the introduction, with the possible exception of competencies, a single feature that included at least two elements (prestige and dominance). Because features were aggregated variables, they are more common at both the text record and culture level (see Figure \@ref(fig:featuressupportplot)).

```{r featuressupportplot, fig.width=10, fig.cap="Support of each feature variable at the levels of text records and cultures. Feature support at the text record was defined as the adjusted proportion of text records with evidence for at least one underlying variable in the feature. Feature support at the culture level was defined as the proportion of cultures with evidence for the feature in at least text record in the culture."}

ggsave('Figures/plot_features_support.pdf', plot_features_support, width = 10, height = 5)
plot_features_support
```

There were modest or no correlation between features, after aggregating leadership dimensions within features (see Figure \@ref(fig:featurecor)).

We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped). As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC. The large increases for group context and leader sex indicated that including these factors substantially improved fit. See Figure \@ref(fig:heatmapdrop1)B. We then compared the levels of each factor that improved fit for each feature variable. Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality. Military groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies. See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).

<!-- ## Exploring population-level variation -->

<!-- To explore potential population variation in the *Prestige vs. Dominance* and *In-group favoritism vs. Social Status* leader qualities components and the *Social functions* and *Mediation vs. Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures. For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).  -->

<!-- For predictor variables we incorporate additional population-level measures. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.  -->

<!-- We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification. Fifty of these cultures are the same as, or very similar to, cultures in the PSF. Specifically, we use the SCCS measures of community size, population density, and cultural complexity. Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders. We also include our researcher coded measure of group structure (See Table X) as a predictor.   -->

<!-- Functions PC 1 model -->

<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
<!-- ``` -->

<!-- Functions PC 2 model -->

<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
<!-- ``` -->

<!-- Qualities PC 1 model -->

<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
<!-- ``` -->

<!-- Qualities PC 2 model  -->

<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
<!-- ``` -->

## Shaman leaders and competencies

```{r shamanstats}
supernatural_n <- sum(leader_text2$qualities_Supernatural==1)
shamanism_pct <- signif(100*sum(df_shaman$shamanism)/nrow(leader_text2), 3)
```

The competencies feature combined dimensions of prestige-style leadership, such as knowledge and experience, with dimensions of dominance-style leadership, such as aggressiveness and feared. It also included supernatural qualities. Shamans and other leaders with supernatural qualities were surprisingly common in the ethnographic record. Of our `r nrow(leader_text2)` text records, `r sum(df_shaman$shaman)` used the word "shaman", `r supernatural_n` provided evidence for supernatural leader qualities, and `r sum(df_shaman$shamanism)` provided evidence for one or both (`r shamanism_pct`% of all text records). We term the latter variable *Shamanism*. We fit a logistic elasticnet model of shamanism as a function of all leadership dimensions (minus *Supernatural*). This model revealed that shaman leaders combined dimensions of dominant leaders (*Feared*) with dimensions of prestigious leaders (*Medical functions*, which are a form of knowledge or expertise, and *Experienced/accomplished*). See Figure \@ref(fig:shamanism).

```{r shamanism, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models of evidence for *Shamanism*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *Shamanism* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *Shamanism*. Both models used the lasso penalty ($\\alpha=1$) and chose $\\lambda_{1se}$ by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 1 not displayed. X-axes are odds ratios on a log 10 scale."}
plot_shamanism_text <- plot_shamanism_text + scale_x_log10()

p <- plot_shamanism_text + plot_elastic_shamanism +
  plot_annotation(
    title = 'Predictors of Shamanism',
    tag_levels = 'A', 
    caption = 'Odds ratio', 
    theme=theme(plot.caption=element_text(hjust=0.5, size=12, margin=margin(t=0)))
    )
ggsave('Figures/shamanism_plot.pdf', p, width = 8, height = 4)
p
```

## Cognitive, social, and somatic capital

Motivated by our feature analysis, and the apparent importance of shaman leaders, we tenatively advance a scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby leaders possess some combination of cognitive capital, social capital, and somatic capital that they deploy to provide benefits and impose costs (for further justification, see the Discussion section). Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]

To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)). We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic elements of leadership among all text records (an 'upset' plot) [@gehlenborg2019]. Horizontal bars (left): The total number of text records in each element. Black dots: has evidence. Gray dots: does not have evidence. Vertical bars (top): The number of text records in that intersection (with lower limit set at 3). Red: The number of text records involving shamans or leaders with supernatural qualities."}

shamanfn <- function(row){
  row['shamanism'] == 'Shaman'
}

p <- upset(
  df_fivefold2,
  nintersects = 18,
  order.by = 'freq',
  point.size = 5, 
  text.scale = 2,
  queries = list(list(query = shamanfn, color = 'red', active=T))
)
pdf("Figures/upset.pdf", width = 12, height = 9)
p
dev.off()
p
```

Figure \@ref(fig:fivefold) reveals that descriptions of leadership in the ethnographic record includes a heterogeneous mix of these five elements. In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs. Shamans are associated with evidence for *Cognitive capital* and imposing costs.

<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019]. Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination. We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.  -->

 <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

# Discussion

This exploratory study systematically evaluated evidence for `r length(all_study_vars)` dimensions of leadership in a stratified random sample of the ethnographic record of `r nrow(leader_cult)` largely nonindustrial (non-WEIRD) cultures that included hunter-gatherers, pastoralists, horticulturalists, and agriculturalists. The quality and function dimensions were *not* determined prior to the study but were instead identified and coded during multiple passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section). Results revealed universal dimensions of human leadership as well as important variation by group context, subsistence type, continental region, and sex of the leader.

## Universal dimensions of leadership

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures, and are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures (see Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductively, or socially in 50% or more of cultures (see Figure \@ref(fig:varsupportcb)). The composite feature measures mirrored these results, finding evidence in 85% of cultures or more, that leaders had social, reproductive, or material success, exhibited a number of competencies, organized and strategized group cooperation, and were prosocial. See Figure \@ref(fig:featuressupportplot).

<!-- Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups...

Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent, are experienced, and have a demonstrated history of previous accomplishments or special capacities. They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.

In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, and military groups; and in political contexts at the community and supracommunity levels. A large percentage of text records in all subsistence groups documented political leadership at the community level. Leadership at the supracommunity level was also common across all subsistence groups save hunter-gatherers, which are largely autonomous. Economic, military, and religious group leadership, although relatively less frequent, were fairly consistently represented across subsistence levels.

Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however (see Figure \@ref(fig:grouptype)).
-->

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous, and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors). See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

<!-- Qualities	Knowledgeable/intelligent	81.4 -->
<!-- Functions	Resolve conflict	78.0 -->
<!-- Functions	Misc. social functions	72.9 -->
<!-- Leader Benefits	Material resources	69.5 -->
<!-- Leader Benefits	Increased social status	66.1 -->
<!-- Qualities	Wealthy	66.1 -->
<!-- Functions	Group representative	64.4 -->
<!-- Qualities	Strategic nepotism	64.4 -->
<!-- Qualities	Generosity	61.0 -->

The leadership features we derived by clustering leadership dimensions were represented across most cultures. Leaders can be conceptualized as performing three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions. Regarding the traits characteristic of leadership, leaders can be conceptualized as possessing some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions. See Figure \@ref(fig:featuressupportplot).

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for the services they provide followers, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.

```{r matingbenefit}
matingbenefit <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable=='Mating benefit'], 3)

reprosuccess <- signif(100*features_support_cult$Estimate[features_support_cult$Variable == 'Social_material_success'], 3)
```

Most potentially universal dimensions discovered here appear in a model developed by @hagen_leadership_2019, in which knowledgeable and intelligent individuals provide computational (cognitive) services to group members in exchange for a variety of fitness benefits, and these services often involve conflict resolution and other social functions [see also @garfield_evolutionary_2019]. The computational services model emphasizes mating benefits, however, which appear in `r matingbenefit`% of cultures, below our arbitrary 60% threshold for "universality" (see Figure \@ref(fig:varsupportcb)). But there was evidence for the *Social, reproductive, and materials success* feature, which included leaders in polygynous unions, having a high quality spouse, and having many children, in more cultures (`r reprosuccess`%) than any other feature (see Figure \@ref(fig:featuressupportplot)). Interestingly, the computational services model highlights women's family leadership role, and the estimated marginal mean for the *Social, reproductive, and materials success* feature was higher for females leaders than male leaders (see Figure \@ref(fig:heatmapfeaturesex)), with the caveat that there were few text records on female leaders in our sample.

Pawnee chiefs illustrate leadership involving several of the universal dimensions, including intelligence, conflict resolution, and generosity [@nq18-015]:

>...a chief was a regulator, not an absolute ruler or tyrant. Although chiefs had considerable authority, their decisions were generally based on a consensus of opinion rather than arbitrary whim. The chief, like his celestial forebear, was supposed to be a guardian of the people, always mindful of their wishes and needs. And even though the office was hereditary in certain families, the man chosen to fill it had to demonstrate humility, generosity, and sagacity, because a jealous or aggressive temperament was considered unbefitting a chief.

Coercive authority is a dimension of leadership that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals. This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992]. Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies. See Figure \@ref(fig:elasticcoercive). This passage describing Trobriand chiefs is illustrative [@oj29-007, pp. 64-65]:

>One point in connection with the chief's authority deserves special mention. Power implies not only the possibility of rewarding, but also the means of punishing. This in the Trobriands is as a rule done indirectly, by means of sorcery. The chief has the best sorcerers of the district always at his beck and call. Of course he also has to reward them when they do him a service. If anyone offends him, or trespasses upon his authority, the chief summons the sorcerer, and orders that the culprit shall die by black magic. And here the chief is powerfully helped in achieving his end by the fact that he can do this openly, so that everybody, and the victim himself knows that a sorcerer is after him. As the natives are very deeply and genuinely afraid of sorcery, the feeling of being hunted, of imagining themselves doomed, is in itself enough to doom them in reality. Only in extreme cases, does a chief inflict direct punishment on a culprit. He has one or two hereditary henchmen, whose duty it is to kill the man who has so deeply offended him, that actual death is the only sufficient punishment.

## Variable dimensions of leadership

```{r miscstats}

movement_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Movement/migration"], 3)*100

moral_pct <- signif(functions_support_cult$Estimate[functions_support_cult$Variable == "Moral authority"], 3)*100

fairness_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Fairness"], 3)*100

feared_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Feared"], 3)*100

attractive_pct <- signif(qualities_support_cult$Estimate[qualities_support_cult$Variable == "Attractive"], 2)*100

movement_hg_pct <- signif(100*sum(all_data$functions_NewSettlement[all_data$subsistence=='hunter gatherers'])/sum(all_data$functions_NewSettlement), 3)

```

There was substantial cross-cultural variation in evidence for most dimensions of leadership: of our `r length(all_study_vars)` dimensions of leadership, `r sum(df_compare2$Estimate < 0.5)` had evidence in less than 50% of our `r nrow(leader_cult)` cultures. See Figures \@ref(fig:varsupport) and \@ref(fig:varsupportcb). There was *systematic* variation in `r vary_n` leadership dimensions by group context, subsistence strategy, continental region, and/or leader sex, factors for which we had complete data for all text records.

These factors were confounded, i.e., group context was associated with subsistence strategy, which was associated with continental region. See Figures \@ref(fig:grouptype) and S\@ref(fig:regionsubsistence). Specifically, evidence for leadership in residential subgroups and kin groups was relatively more common in hunter-gatherer and pastoralist societies, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013; @venkataraman_hunter-gatherer_2017]. Evidence for supracommunity political leadership was more common in horticultural and agricultural societies, which tend to rely heavily on cultivated land and maintain fixed, defensible territories. These patterns support perspectives suggesting residential mobility and increased costs of dispersal, subsistence intensification, and demographic transitions are associated with increasing complexity of sociopolitical organization among non-WEIRD populations [@johnson_evolution_1987; @steward_basin-plateau_1938; @powers_evolutionary_2014; @kantner_identifying_2010]. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese by @az02-001[p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).

Pastoralists, despite potential residential mobility, do rely in part on a highly defensible resource base (i.e., livestock) and exhibit evidence for political leadership beyond the community more so than hunter-gatherers and on par with horticulturalists. Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992; @glowacki_evolutionary_2017]. For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p. 58] notes that, "the Sultan does not lead the army of the tribe in war. He names a chief of the armed forces."

One possible interpretation of the negative relationships between evidence for political leadership at the community level or above and leadership in kin groups or residential groups is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes. It is interesting, however, that even among nonindustrial agricultural populations which are characterized by political leadership systems beyond the community level -- more similar to political structures of WEIRD populations -- the evidence of leadership within residential and family groups persists. For example, among rural Korean populations a chairwomen of local 'club' organizations maintains an important, although informal, leadership position within the group and represents the group to governmental bodies. @aa01-062 [p. 173] describes,

>the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity. In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process. She is not a figurehead but rather a middleman in the meeting, functioning, in effect, as a go-between for the group. There is no despotic idea in the context of a total consensus in this essentially egalitarian society. The chairwoman stands as a middleman between the government and the village. Although the government may call her a ‘leader’ in the village, she properly does not occupy this role.

Because group context, subsistence strategy, and continental region were confounded (and there was some evidence that leader sex was too), we fit multiple regression models with all four factors included. Interestingly, the `r vary_n`  systematically varying dimensions of leadership tended to be associated with one factor above the others. See Figure \@ref(fig:heatmapdrop1).

The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership. See Figure \@ref(fig:heatmapgroups). These results underscore recent trends in evolutionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

Four leadership dimensions varied by subsistence strategy. There was more evidence that hunter-gatherer leaders lacked coercive authority compared to other leaders, which supports much theory and ethnography discussing egalitarianism and the active resistance of hierarchy among hunter-gatherers [e.g., @boehm_hierarchy_1999;@boehm_purposive_2008; @woodburn_egalitarian_1982]. Reliance on a mobile and stochastic resource base is suggested to promote strong social norms of sharing [@peterson_demand_1993; @cashdan_egalitarianism_1980] and shape both resource transfers and partner preferences [c.f., @smith_friend_2018; @smith_hunter-gatherers_2018]. There was more evidence that both hunter-gatherer and pastoralist leaders generated food surpluses, but less evidence that pastoralist leaders distributed food and other resources. Thus, systematic ethnographic evidence supports the important role of provisioning resources by influential males among foragers [@wood_prestige_2006; @gurven_give_2005; @gurven_men_2009; @alger_paternal_2020]. However, there was more evidence that leaders in communities with mixed subsistence strategies distributed resources, which lends support to males using economic success and productivity as an influence-seeking strategy [@hawkes_showing_2002; @hawkes_family_2010].

Six leadership dimensions varied by continental region (see Figure \@ref(fig:heatmapregion)). We do not have clear theoretical interpretations of this variation, so we treat region simply as a control variable.

## The rarity of some important leadership dimensions

Evidence for several theoretically important leadership dimensions was relatively rare in the ethnographic record. The nonhuman animal leadership literature, for example, commonly focuses on group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among small-scale human societies and nonhuman animal societies, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (`r movement_hg_pct`%) were hunter-gatherers. Among the Blackfoot in the North American plains for example, @nf06-061 [p. 411] describes that, "if there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south. He did not order the others to follow, but he knew that they probably would go with him."

Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet evidence for leaders as sources of moral authority or "fair" was found in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively). The dominance theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures was there evidence that leaders were feared. Finally, a large literature on leadership in industrialized societies highlights the physical attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but evidence that leaders were attractive was found in only `r attractive_pct`% of the nonindustrial cultures sampled here. Clearly there is substantial diversity in many dimensions of leadership among non-WEIRD populations.

<!-- WARFARE H&G Figure S\@ref(fig:heatmapsubsis).  -->

## Benefits and costs of leadership and followership

The evolution of leadership and followership is somewhat perplexing. Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs. Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest. The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained [e.g., @gavrilets_solution_2014; @hooper_theory_2010; @price_evolution_2014; @bastardoz_nature_2018; @hagen_leadership_2019; @garfield_review_2019].

We assessed the benefits and costs of leadership in two ways. First, we explicitly coded each text record for evidence of benefits and costs. Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high-risk, high-return strategy (see Figure \@ref(fig:varsupportcb)). Nevertheless, for most dimensions of benefits and costs, there was more evidence of benefits than costs (Figure \@ref(fig:benefitcostratio)), and the same was true averaging across all benefit/cost dimensions ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$). Second, several positive leader qualities indicating social, reproductive, and material success clustered together (bottom cluster Figure \@ref(fig:cluster)A). Both results, while not providing evidence of net fitness benefits, are nevertheless encouraging, especially since leaders had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and also had more evidence for inclusive fitness, material and social benefits, relative to costs. This is consistent with much of the theoretical literature that links leader benefits to their capacities to provide prosocial benefits (e.g., resolving conflicts, organizing collective actions, and being generous) and/or impose social costs (e.g., punishing and aggressiveness) [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013; @hooper_theory_2010; @gavrilets_solution_2014; @price_evolution_2014; @smith_leadership_2016].

Among the Trobriand horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts: 
>Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy. Rank was quite directly expressed in the number of wives. The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19). In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136). In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists. An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....

The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Followers were not reported to receive mating benefits, relative to leaders or other benefit types. For both leaders and followers, individuals stood to lose what they were able to gain. See Figure \@ref(fig:varsupportcb). @nm09-056 [p. 183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois, whose life exemplified the risks to both leaders and followers:

>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century. It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment. Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades. Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period. Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.

## Disentangling the relationship between leadership and high status

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019]. Drawing on a comparative-evolutionary perspective, @van_vugt_dual_2019 emphasize that although leadership and high social status do often covary, leaders are not always high-status individuals, exemplified by low-ranking individuals in nonhuman species that lead group movement or younger and lower status men and women who initiate collective actions. @cheng_dominance_2019 emphasizes, however, that among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership [see also @cheng_why_2020]. The "low status" young adults who initiate a collective action might be high status among their peers, within what @garfield_review_2019 [p. 71] described as the "intragenerational social hierarchy."

In our view, social context is critical to these discussions [see @wiessner_power_2010]. Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who organize the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012]. @garfield_evolutionary_2019 found that leader organization variables clustered separately from leader prestige variables, suggesting they are distinct. @garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect. Some older individuals, on the other hand, were highly respected but had little social influence or authority.

In our results, high status was the most common quality of leaders at both the text record  (`r statusPercTxt`%) and culture level (`r statusPercCult`%) (Figure \@ref(fig:varsupport)); it was also the second most common benefit of leadership and the loss of status was the most frequent cost (Figure \@ref(fig:varsupportcb)).
<!-- mentioned in this sample of the ethnographic record which includes discussions of chiefs, work group leaders, shamans, and age-sets, from a culturally diverse sample including foraging groups and kingdoms.  -->
This lends support to the notion of high social status often being a necessary condition of leader emergence [@cheng_why_2020]. Nevertheless, over two thirds of our text records on leadership did not mention status, and it was not one of our candidate universal dimensions of leadership. Interestingly, evidence for high status leaders varied by continental region, with more evidence for them from South American cultures (see Figure S\@ref(fig:heatmapregion)).

The dimensions of leadership predictive of high status leaders (see Figure \@ref(fig:highstatusplot)B) suggest the phenotype of such leaders resembles what many anthropologists refer to as "Big men" [@garfield_review_2019]: wealthy, accomplished older men who favor members of their kin group, and are highly regarded for their decision-making abilities, knowledge, and intelligence. High status leaders are able to direct benefits to their close relatives, likely due, in part, to their greater material and social wealth, which implicates leadership dynamics as a driver of the evolution and emergence of inequality [c.f., @sanday_female_1981; @shenk_status_2016; @price_social_2010; @borgerhoff_mulder_pastoralism_2010; @mattison_evolution_2016; @bowles_emergence_2010]. Our results did not strongly implicate high status leaders with particular functions, with the exception that punishment was modestly negatively associated with evidence for high status.

@smith_leadership_2020 speculate that, "leadership may...be more strongly correlated with high status in large, complex organizations, such as in corporations and governments (p. 2)." Evidence for high status leaders, however, did not substantially vary by group structure, which included levels of social organization ranging from sub-residential to supracommunity (see Figure S\@ref(fig:groupstatus)). This null result is evidence against the @smith_leadership_2020 hypothesis.

Overall, our results suggest that although high status leaders are common across cultures and all levels of social organization, only some leaders are high status, namely, those with a limited set of qualities such as wealth, experience, decision-making abilities, intelligence, and interpersonal skills, as well as older age and conformity to social norms.

## Female leadership and sex differences

<!-- CITE OF CVR EMPIRICAL PAPER NEED TO DO -->

Only `r sum(all_data$demo_sex == 'female')` of the `r nrow(all_data)` text records discussed female leaders exclusively, and another `r sum(all_data$demo_sex == 'both')` explicitly discussed both female and male leaders. Leader sex was only associated with variation in seven of our `r length(all_study_vars)` dimensions of leadership (see Figure S\@ref(fig:heatmapsex)), and one feature variable (see Figure S\@ref(fig:heatmapfeaturesex)), indicating that most dimensions of leadership probably characterize leaders of both sexes.

When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)). Female leaders also had higher levels of evidence for the social, reproductive, and material success "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)). Evidence for female leaders was also more likely to be found in kin groups and residential subgroups. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004]. @garfield_investigating_2019 similarly found that there was positive assortative mating of high status leaders among elected leaders in a small-scale society with relative gender-egalitarianism. Overall, our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].

## Beyond dual models of leadership

Leadership and social influence are expected to be associated with particular suites of individual qualities and functions, which potentially covary by cultural or social context, personality, or other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001; @bass_handbook_1990]. Our systematic cross-cultural results support many aspects of influential evolutionary theories of leadership, but offer new direction for the expansion and synthesis of leadership theories.

**Dominance and prestige**. Much of the evolutionary theoretical literature on leadership has suggested social influence is most often achieved via one of two distinct strategies: dominance or prestige [for review see @garfield_review_2019; @garfield_evolutionary_2019]. There is a long history of leadership styles being described as based on either force and strength (i.e., dominance) or respect and special skills (i.e., prestige). Such dual-models are rooted in the coupling of ethology and animal dominance hierarchies with ethnography and cultural prestige systems [see @mead_sex_1935; @sahlins_poor_1963; @lewis_leaders_1974; @tiger_dominance_1970; @bernstein_dominance_1976; @barkow_prestige_1975]. Synthesizing the ethnographic and ethological evidence, @kracke_force_1978 linked "force and persuasion" to psychological-affect between leaders and followers in small-scale societies and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] suggested the evolution of human prestige was linked to intra- and inter-sexual selection of culturally acquired competencies.

@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on the evolutionary roots of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies. @henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019]. @van_vugt_dual_2019 [p. 1], for example, "argue that there are two leadership styles: prestige style and dominance style," and @cheng_two_2013 [p. 119] in focusing on distinctions between dominance and prestige suggests, "human social stratification is characterized by the co-occurrence of both strategies." Our analyses support the cross-cultural importance of leader aggression and coercion, as well as culturally acquired competencies, but they also indicate that some important modifiications are required.

<!-- DISCUSSES COMPETENCIES FEATURE -->
**Leader competencies and shamanism**. The individual competencies cluster (Figure \@ref(fig:cluster)A) comprises variables associated with dominance-based leadership, including coercive authority and aggressiveness, as well as qualities associated with prestige-based leadership, including knowledgeable/intelligent and experienced/accomplished. These dimensions are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)).

Shamanism appears to be a distinct form of leadership that combines a strategy of inducing fear, similar to the dominance strategy, but is based on knowledge and expertise, similar to the prestige strategy. The *Competencies* feature included supernatural and feared dimensions (Figure \@ref(fig:cluster)). Our *Shamanism* variable (which was not one of our original leadership dimensions) was predicted by words that evoke both dominance and prestige, such as *supernatural*, *power*, *fear*, and *knowledge* (Figure \@ref(fig:shamanism)A), and by four other leadership dimensions that also overlap with the dominance and prestige constructs: *Feared*, *Medicinal functions*, *Ritual functions*, and *Experience/accomplished*. @winkelman1987cross found that in societies sampled from the HRAF ($n = 43$) shamans provided decision-making services within judiciary, economic, military, or political domains in 44% or more of societies, and held charismatic leadership roles in 19% of societies; @winkelman1992shamans emphasizes the influential positions of leadership shamans attain through their charisma, social unification, healing abilities, and use of supernatural powers to cause harm. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate superhuman abilities [@singh_cultural_2017].

@ng06-054 describes shaman leaders among the Ojibwa:

>Leadership was vested in the heads of the co-residential groups who held their position through their hunting abilities and supernatural power as shamans. Shamans had the ability to foresee future events and the ability within certain limits to control them. These men were feared as well as respected and prior to 1900, most leaders were polygynists. Leadership in 1900, was based upon individual initiative and the vision quest and was not inherited, a fact that is congruent with the bilateral structure of Ojibwa society at this date (cf. Helm 1965).

<!-- As brief examples of the diversity of leader capacities, of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p. 85]; and among the Highland Scots, @es10-003 [p. 147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy. -->

<!-- DISCUSSES BOTH MANAGEMENT FEATURES -->
**Leaders are managers**. There is a history within managerial studies among WEIRD populations of contrasting *leaders* from *managers*. @zaleznik_managers_1977, for example, suggested "managers" seek systematicity, organization, and stability, whereas "leaders" are willing to accept a lack of control and creatively navigate uncertainty to produce novel insight and appeal to the emotions of followers. Political anthropology reveals, however, in the absence of state-based or large-scale institutions, local leaders are often tasked with both managing various economic systems while also providing social services and interacting with followers on an interpersonal level [@cohen_comparative_1967; @fried_evolution_1967]. Within the domain of leader management, there are likely to be two primary features of human leader functions: strategizing and organizing. Our results (Figure \@ref(fig:cluster)B) correspond very closely to what @van_vugt_cognitive_2007 define as *strategic leadership*, where leaders increase the pay-offs of another individual's behavior thereby incentivizing followership, versus *coordinating leadership*, where individuals bound by group challenges must effectively organize collective effort.

Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980]. Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p. 185] describes the nature of conflict resolution processes in the absence of strong authority:

>The elder settles numerous disputes, both within and without his hunting band. Such disputes mainly concern marital disagreements, Schebesta says. The elders often settle controversies by delivering speeches during the evening, leading frequently to noisy debates between the elders of different bands or lineages. Outside the lineage context, the elder is not in any way different from the others; there are none of the trappings of villager chieftainship. He delivers his decisions without hesitancy and is obeyed in the same way within his lineage. His power is limited, however, by custom and tradition as well as by context. He may neither ignore nor alter traditional practices.

The feature of leader organizational functions aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p. 127] notes:

>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual. A mahsa wami must hold sway over all three areas to lead effectively. Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor. Control, then, implies the obligation to redistribute. The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict. I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.

<!-- PROSCIAL FUNCTIONS FEATURE -->
<!-- DISCUSSES PROSOCIALITY FEATURE -->
**Prosociality**. The prosocial cluster (Figure \@ref(fig:cluster)A), comprising prosocial competencies such as generosity, interpersonal and oratory skills, and charisma, and cultural conformity, such as proper behavior and having a favorable personality, generally corresponds to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @grabo_charisma_2017; @howell_role_2005]. The importance of leader prosociality also supports the @henrich_big_2015 Big-man mechanism by which the positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership.
<!-- A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical formidable*, and *Drug consumption* (although drug use is often an important social activity). -->

We also found that leader functions often involve prosocial investments in group welfare (Figure \@ref(fig:cluster)B). The prosociality feature of leader functions included distributing resources, providing protection, and various prosocial investments, which are functions commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987]. In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain their social influence via continued community investment and the demands of followers can shape the functions of leaders to meet local needs.

## Leadership and cognitive, social, and somatic capital

In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers". These models, in short, have conflated the ability to provide "benefits" with knowledge and expertise, and the ability to impose "costs" with physical formidability. @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1). Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability. The use of knowledge to impose costs is widely recognized by many social scientists. The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006] and also among hunter-gatherers [@hess_interpersonal_2010].

If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019].

# Limitations

The ethnographic record is an imperfect source of information on human cultural diversity. Our results are conditional upon the content ethnographers chose to discuss, the way they chose to discuss it, and the way we interpreted and coded their texts. There was a general male bias, which could reflect a male bias in community leadership or a more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988]. These data are also biased towards cultural models of leadership (rather than particular cases of leadership) and community leadership; descriptions of kin group and residential group leadership were less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally. Our sample of cultures and text records by continental region and subsistence strategy is also unbalanced. We also detected bias in evidence by date of publication for nine of our 109 leader dimensions (see Figure S\@ref(fig:pubdate)).

Another limitation lies in our ability to interpret the absence of evidence. We are confident that the cross-culturally frequent measures discussed here are strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns. We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.

This is an exploratory study, which means that our analyses were driven by the data rather than by *a priori* hypotheses. This has the advantage that unexpected patterns can be discovered, but the disadvantage that some random variation in the data can, and almost certainly will, be misinterpreted as meaningful patterns. In particular, our cluster analysis of leader qualities (Figure \@ref(fig:cluster)A) was sensitive to the choice of different clustering algorithms and inclusion or exclusion of variables with little evidence.

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

<!-- OUR SUPPORT FOR THEORIES -->
Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs. We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

<!-- SHAMANS AND BEYOND DUAL MODEL -->
Shamans emerged as an important category of leaders that are rarely discussed in the evolutionary literature [but see @singh_cultural_2017]. Supernatural qualities were among the top ten leader qualities across cultures (Figure \@ref(fig:varsupport)). Shamans share qualities with prestige-style leaders, such as having specialized knowledge and abilities, yet appear, at least at times, to use their knowledge to instill fear, similar to dominance-style leaders. We therefore suggest moving beyond dichotomous or "dual" models of leadership, which have tended to conflate knowledge and expertise with providing benefits. Instead, our data indicate that leaders employ cognitive, social, and somatic capital to provide benefits and impose costs (Figure \@ref(fig:fivefold)). This view aligns with the theoretical framework developed by many scholars that emphasizes the importance of embodied capital (neural and somatic) and social and material capital in the evolution of human social organization [@kaplan2003neural; @kaplan_embodied_2003; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @bowles_emergence_2010].

<!-- and focusing on inter-individual variation in broader typologies of leadership competencies. There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies. The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies. Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer. -->

<!-- DIVERSE FUNCTIONS -->
<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019]. -->

<!-- Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.  -->

<!-- LEADER SUCCESS AND INEQUALITY -- REPETITIVE ? -->
<!-- The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership. -->

<!-- CONNECT TO WERID -->
Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriented* framework of managerial behaviors (i.e., our Management vs. Prosociality features of leader functions). However, our results, which found evidence for oratory skills in about 50% of cultures and charisma in about 20%, provide only partial support for the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic]. See Figure \@ref(fig:cluster).

<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].  -->

<!-- CLOSING STATEMENT -->
Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics. Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @smith_cultural_2020; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mattison_evolution_2016].

# Supplementary information

## WEIRD leadership dimensions

Below we briefly review leadership functions and qualities in WEIRD contexts to suppliment our discussions of non-WEIRD leaderhsip dynamics in the main text.

**WEIRD leadership functions**. In WEIRD societies, community and political leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions. The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications. Within local communities and in smaller-scale economic groups, however, leaders in WEIRD populations *are* likely to be direlty involved in mediating conflicts and managing local cooperative projects [e.g., @acheson1988lobster; @rich_dynamics_1980].

<!-- THEORY ON WEIRD FUNCTIONS -->
Scholars in managerial studies have developed multiple models on the functions of leadership. Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990]. These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics. Later theorists incorporated cognitive and socioemotional components of leaders functions. For example, @selznick_leadership_1957 describes the functions of organization leaders are to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.

<!-- DATA ON WEIRD FUNCTIONS -->
@wofford_factor_1970 provides an early empirical view of managerial functions in a WEIRD setting. A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions. Wofford's analysis provided a foundation for future empirical work on management. He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].

<!-- COMPARATIVE DATA ON WEIRD FUNCTIONS -->
More recently, some managerial scholars have adopted comparative methods. @behrendt_integrative_2017 suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources. This framework is intended to provide a more externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics. 
<!-- More integrated models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.  -->
Culturally Endorsed Implicit Leadership Theory (CLT), suggests social norms impact follower preferences for leadership styles [@moore_literature_2018]. The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations. The GLOBE project has produced multiple cross-national data sets on and hundreds of publications. *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* [@house2013strategic] uses data from over 1,000 CEOs and over 5,000 senior executives from varied corporations across 24 countries to address (1) national cultural influences on the behaviors of leaders, (2) the behaviors of CEOs typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations. The study concluded cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations. That is, @house2013strategic suggest leaders moderate their behavior through culturally learned values and leadership-specific values.

The GLOBE CEO Study defined six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, they found charismatic leadership behavior was consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and humane-oriented behavior. 
<!-- Participative leadership was moderately related to TMT dedication, however was not related to firm performance.  -->
The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive. Furthermore, results strongly suggest the fit between a CEO's behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance. By developing higher-order features of leadership behavior and situating leadership functions within more detailed frameworks, insights from managerial studies can more effectively be coupled with approaches from across behavioral science.

**WEIRD leadership qualities**. Drawing on cross-national data through the GLOBE project from over 15,000 middle managers in businesses across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership. In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.

<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->

Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002]. More complex organizations are thought to require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view among diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009]. Some recent evidence, however, suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.

**WEIRD leadership costs and benefits**. Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

## Ethnographic sample

The 60-culture Probability Sample Files of the HRAF developed by @naroll_proposed_1967 is based on culturally and geographically diverse populations which are well covered in the ethnographic record. The distribution of cultures, however, is not equal across categories of subsistence strategies; agriculturalists and horticulturalists are relatively over-represented compared to hunter-gatherers and pastoralists. The ethnographic sample of text developed in the *leadershipdata* package is also not equally balanced across all subsistence strategy categories. See Figure S\@ref(fig:subsisttree).

```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012). The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and subsistence strategy."}

plot_cult_docs_subsis
```

```{r}
doc_1950_perc <- signif(length(documents$d_publication_date[documents$d_publication_date>1949])/length(documents$d_publication_date),2)*100
```

The majority of ethnographic documents sampled here (`r doc_1950_perc`%) were published after 1950 and the mean publication year is `r round(mean(documents$d_publication_date),)` (SD = `r round(sd(documents$d_publication_date),2)`). Figure \@ref(fig:doc-years) plots the publication years of all documents which contributed text records to the ethnographic sample.

```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the median publication year."}
doc_year_hist
```

The eHRAF provides information on the time frame of field work associated with each document. For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field. For most documents one or more spans of years are given. Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document. The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred. We omitted data for one document from North America with field work date ranges prior to 1805. For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

```{r field-years, fig.height=8, fig.width=9, fig.cap="Span of fieldwork years of study documents. To improve clarity, one document with very early fieldwork from 1791-1804 was omitted."}
doc_fielddates_plot
```

## Continental region is confounded with subsistence strategy

The distribution of cultures by subsistence strategy across continental region is also variable. In this sample, most African cultures are agriculturalists and most North American cultures are hunter-gatherers, for example. See Figure S\@ref(fig:regionsubsistence).

```{r regionsubsistence, fig.width=12, fig.cap="Proportion of text records for different regions and subsistence types. Size of rectangles is proportional to the number of text records."}

lbls <- c(
  'Africa', 
  'Circum-\nMediterranean', 
  'East\nEurasia', 
  'Insular\nPacific',
  'North\nAmerica',
  'South\nAmerica'
  )

ggplot(all_data) + 
  geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) +
  scale_fill_viridis(discrete = T) + 
  scale_x_productlist(labels = lbls) +
  guides(fill = guide_legend(reverse = T)) +
  labs(x = '', y = '') +
  theme_minimal(15)
```

## Bias

```{r eHRAFpages, fig.height=8, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF. Each dot is one culture. Line fit by loess regression."}
plot_pages_tr
```

We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs. male-only authorship, and variation in the total number of pages of ethnography for each culture. For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures. We fit two sets of models. The first univariate set fit separate models for each study variable and each bias variable (total number of models = $`r length(all_study_vars)` \times 3 = `r length(all_study_vars)*3`$). The second multivariate set included all bias variables in each model (total number of models = `r length(all_study_vars)`). All p-values were corrected for a FDR of 5%. After correction, the coefficients of female authorship and pages of ethnography were not significant in any model. Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model. See Figure \@ref(fig:pubdate).

```{r pubdate, fig.cap=glue_data(list(n=length(all_study_vars)), "The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs. Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture. All models had random intercepts for ethnographers nested within cultures. Models were fit for each of the {n} variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant. Publication date was centered and standardized by 1 SD. Each dot is one coefficient (converted from log odds to odds ratios). Bars are 95% CI.")}
plot_pubdate
```

## Predicting high status leaders with our four factors

<!-- These data present an opportunity to explore the relationship between high status and leadership, as represented in ethnographic texts. To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words). In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record. The coding of *high status* was joined with this term-document matrix. This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).  -->

<!-- We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is large relative to the number of cases. Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting. The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005]. We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality. We used tenfold cross-validation to choose a $\lambda$ parameter that was within in standard error of the value that minimized the deviance. We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).  -->

```{r highstatusplot2, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda. Red values are positive predictors; purple values are negative predictors. X-axis is on a log 10 scale."}
# highstatus_plot + scale_x_log10()
```

<!-- Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficients. The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*. -->

<!-- NEED TO DO TABLE -->

As reported in the main text, we also investigated if the proportion of text records with evidence for *High status* differed among the different levels of group context, subsistence strategies, continental regions, and leader sex (again using a logistic mixed effects model with these variables as fixed effects and authors nested within cultures as random intercepts). There were no clear differences and we interpret this as a null result. See Figure S\@ref(fig:groupstatus).

<!-- Hunter-gatherers and pastoralists were more likely to provide evidence for high status leaders, as were South America and East Eurasia. Female leaders were also more likely to be identified as high status. Somewhat counter-intuitively, the proportion of text records indicating high status leaders was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status. See Figure \@ref(fig:groupstatus). -->

```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A. subsistence strategy, B. group structure, C. continental region, and D. leader sex. Model included random intercepts for ethnographer nested within culture. Values are on the response scale (probability)."}
# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')

theplots <- all_emms(m_status_group2, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.8, 'Predictors of leader status')

((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader status', tag_levels = 'A')
```

## Relative effects of subsistence strategy, region, leader sex, and group structure on leadership dimension

Below we plot the estimated margianl means of the effects for level of evidence for each of the `r vary_n` (out of `r nrow(multi_allvars)` total) which produced a model that showed an improvement *with* the four factor predictors over an intercepet only model, by the variable which produced the model improvement. See Figure \@ref(fig:heatmapdrop1)A in the main text. Figure \@ref(fig:heatmapgroups) plots the ledaerhsip dimensions which varied by group structure, Figure \@ref(fig:heatmapsubsis) plots the leadership dimensions which varied by subsistence stragegy, Figure \@ref(fig:heatmapregion) plots the leadership dimnenbsions which varied by contintental region, and Figure \@ref(fig:heatmapsex) plots the leadership dimensions which variaed by leader sex. Table \@ref(tab:variablevars) displays the nbumber of text records providing evidence for each of these the `r vary_n` varying leaderhsip dimennsions and the results of logistic mexed effects models including the AIC scores for the intercept only models, AIC values for models with the four predictors, the AIC difference, and adjusted p-values for the four factor coefficents. Table \@ref(tab:universalvars) displays the same information for the remaining leaderhsip dimensions which did not substantially vary by our four factors.

```{r heatmapgroups, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

p_heatmap_groups + theme_minimal(15)
```

```{r heatmapsubsis, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy. Estimated marginal means were standardized by rows, and therefore can only be compared within rows, i.e., the colors represent the relative amount of evidence across levels of the categorical predictor within the same outcome variable."}

p_heatmap_subsis + theme_minimal(15)
```

```{r heatmapregion, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

p_heatmap_region + theme_minimal(15)
```

```{r heatmapsex, fig.width=10, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

p_heatmap_sex + theme_minimal(15)
```

Note that very few text records (N=`r sum(all_data$demo_sex=='both')`) explicitly mentioned both male and female leaders. Hence, although not depicted here, the confidence intervals for the estimated marginal means for this level are wide. See Table \@ref(tab:sexsubtable) for the distribution of leader sex data by subsistence strategy.

```{r sexsubtable}
kable(sub_sex_tbl, 
      caption = 'Distribution of text records by leader sex and subsistence strategy.')
```

```{r multiout}
multi_out <-
  multi_allvars %>% 
  select(
    var,
    Variable, 
    Evidence, 
    interceptAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_sex,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable)

```

```{r variablevars}
multi_out %>% 
  dplyr::filter(AIC_diff < -2) %>% 
  kable(
  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures. Models were retained if their AIC values decreased by two or more relative to an intercept-only model. Evidence: the number of rows with evidence for the variable. interceptAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each predictor variable in the model.',
  digits = 2
  )
```

```{r universalvars}
multi_out %>%
  dplyr::filter(AIC_diff >= -2) %>% 
  kable(
  caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model. These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence. Evidence: the number of rows with evidence for the variable. interceptAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each predictor variable in the model.',
  digits = 2
  )

```

## Total leader benefits and costs

To supplimnet analyses of costs and benefits we explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, by the predictor variables used above (subsistence strategy, continental region, leader sex, and group structure). There was less evidence for leader costs, however, which limited our ability to fit exploratory regression models. Therefore, in modeling total leader costs, we used subsistence and group structure as predictors. In general, leaders at the highest supracommunity political level had the most evidence for receiving benefits and the least evidence for suffering costs and leaders of military groups has the highest total costs. See Figure \@ref(fig:leadertotalbenefits).

<!-- NEED TO DO -- ADD TABLE? -->

```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A. subsistence strategy, B. group structure, C. continental region, and D. leader sex. Model included random intercepts for ethnographer nested within culture. Values are on the response scale."}
# p1 <- 
#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader benefits", x = "", y = "")
# 
# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader costs", x = "\nProbability", y = "")
# 
# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

theplots <- all_emms(m_ldrtotben, c('subsistence', 'group.structure2', 'region', 'demo_sex'), 0.1, 'Predictors of leader benefits')

((theplots[[1]] / theplots[[2]]) | (theplots[[3]] / theplots[[4]])) + plot_annotation(title = 'Predictors of leader benefits', tag_levels = 'A')

```

```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A. subsistence strategy, and B. group structure. Model included random intercepts for ethnographer nested within culture. Values are on the response scale."}
theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')

theplots[[1]] / theplots[[2]] + plot_annotation(title = 'Predictors of leader costs', tag_levels = 'A')

```

## Dimension reduction

We explored logistic PCA to reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1) using the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(df_qual)` text records for the leader quality variables and `r nrow(df_func)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.

```{r logisticpcafunc, fig.height=8, fig.width=10, fig.cap="Logistic PCA of leader function variables with two principal components."}
logisticPCA_loadings_plot(m_lpca_func, df_func)
```

We then explored associations of PC1 and PC2 with the following culture-level and document level variables: continental region, subsistence strategy, cultural complexity, settlement fixity, population density, community size, discussion of female leadership, female ethnographer, number of pages of ethnography, and publication date. We did not find any convincing associations with any of these culture-level or document-level variables.

```{r lpcafuncbiplot, fig.cap="Function variable PC scores for each text record. Colors represent text records that had evidence for a function variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables). Mixed: evidence for two or more of high-loading variables."}
plot_lpca_func
```

```{r logisticpcaqual, fig.height=10, fig.width=10, fig.cap="Logistic PCA of leader quality variables with two principal components."}
logisticPCA_loadings_plot(m_lpca_qual, df_qual)
```

```{r lpcaqualbiplot, fig.cap="Quality variable PC scores for each text record. Colors represent text records that had evidence for a quality variable that loaded highly on either PC1 or PC2 (but no evidence for the other high-loading variables). Mixed: evidence for two or more of high-loading variables."}
plot_lpca_qual
```

## Feature models

We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`. The Strategize, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/reproductive/material success. Prosocial competencies was correlated with Cultural conformity. See Figure \@ref(fig:featurecor).

```{r featurecor, fig.cap="Feature variable correlation matrix. Values are correlation coefficients."}
plot_feature_cor
```

```{r heatmapfeaturesubsis, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by subsistence strategy. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
# plot(p_heatmap_feature_subsis)
```

```{r heatmapfeatureregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by continental region. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
# plot(p_heatmap_feature_region)
```

```{r heatmapfeaturesex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by leader sex. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
heatmap_feature_sex
```

```{r heatmapfeaturegroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each feature variable by group context. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}
heatmap_feature_groups
```

```{r featurecoefs}

feature_models %>%
  unnest(Tidy) %>% 
  select(
    Variable,
    interceptAIC, 
    multiAIC, 
    AIC_diff,
    effect,
    group,
    term,
    estimate,
    std.error,
    conf.low,
    conf.high,
    statistic,
    p.value
  ) %>%
  group_by(Variable) %>% 
  mutate(
    interceptAIC = c(interceptAIC[1], rep(NA, n() - 1)),
    multiAIC = c(multiAIC[1], rep(NA, n() - 1)),
    AIC_diff = c(AIC_diff[1], rep(NA, n() - 1)),
  ) %>% 
  ungroup %>% 
  arrange(Variable) %>% 
  gt %>% 
  fmt_missing(columns = everything(), missing_text = '') %>% 
  fmt_number(vars(estimate, std.error, statistic, conf.low, conf.high)) %>% 
  fmt_scientific(vars(p.value), decimals = 1) %>% 
  tab_options(table.font.size=10, data_row.padding = 0)

```

## Ethnographer and Culture random effects

```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
plot_fun_auth_cult_re
```

```{r QualityRE, fig.cap="Ethnographer and Culture random effects for quality variables."}
plot_qual_auth_cult_re
```

## Leader market variable operationalization

```{r leadermarket}

frmt <- function(label, vars){
  c(label, rep('', length(vars) - 1))
}

marketvars <-
  tibble(
    Element = c(
      frmt('Cognitive capital', cognitive_vars),
      frmt('Social capital', social_vars), 
      frmt('Somatic capital', physical_vars),
      frmt('Provide benefits', benefit_ability_vars),
      frmt('Impose costs', cost_ability_vars)
    ),
    `Component variables` = c(
    names(cognitive_vars),
    names(social_vars),
    names(physical_vars),
    names(benefit_ability_vars),
    names(cost_ability_vars)
    )
  )
kable(marketvars, caption = 'Operationalization of our leadership components')
```

## GLMM coefficients

```{r glmmcoefs }
# kable(all_coefs)

all_coefs %>% 
  gt %>% 
  fmt_missing(columns = everything(), missing_text = '') %>% 
  fmt_number(vars(estimate, std.error, statistic, conf.low, conf.high)) %>% 
  fmt_scientific(vars(p.value), decimals = 1) %>% 
  tab_options(table.font.size=10, data_row.padding = 0)

```

# Notes

## Possible leader quotes (for universal dimensions)

>This headman caters to the needs of the members of smaller group. The system that such heads are always necessarily recognised at wider level, and that, to coordinate, [Page 85] they have a further head over them does not seem to be in existence. The headman, in fact, is primarily meant to look after petty affairs concerning intra-group life; but if need be, he can be used for resolving inter-group disputes. The position of such heads is neither hereditary nor acquired through formal elections. To recognise a head is something spontaneous, especially when some attention-drawing incidents occur. Age factor is not very vital for headmanship. The man who is physically sound and considered active and witty is picked up. The members of the group start considering his views and regard the same as more than those given by any body else. The headman's views carry weight and are attended accordingly.

>At this point, it is appropriate to sum up the role and typical career of a local leader among the Makuna. The political leader ( [unknown]uh[unknown]u ) is, in general, a man who thinks and speaks well. Oratorical skill is one of the basic requirements and [Page 88] qualities of a prestigious leader. Thinking and speaking well mean that he knows how to deal with people, how to prevent social conflict and how to settle disputes and restore social order. The handling of the potentially disruptive, affinal relationships and the sponsoring of communal rituals are the two most important functions of the Makuna leader. Through his oratorical skill-displayed in ceremonial dialogues-disputes are settled and the interests of his followers defended. Through the communal rituals, the social-and ultimately cosmi—corder is maintained or restored.

>The role of the political leader is well defined in Makuna society. It is a specialist role of the same kind as the ritual specialist roles described above. In fact, the role of the political leader is a direct complement to the role of the shaman: while the shaman is a supernatural mediator-a mediator between men and spirits-the political leader is a social mediator-a mediator between men and social groups. The leader represents his group in relation to outsiders of his own kind (human beings), while the shaman deals with non-human “outsiders”, that is, spirits.
>An informant nearly summarized the typical career of a political leader: “A man who wants to become an [unknown]uh[unknown]u must work hard. He must build a large house, cut large gardens and sponsor many dances. He must make a lot of beer and coca and invite many people. He must think good for his people and know how to greet visitors and how to handle enemies. He must be strong and wise.”

>Such headmen, especially the privileged ones, compete vigorously against each other for renown. The privileged headmen or chiefs are loosely inter-graded amongst themselves in ascriptive prestige terms. This fluid scaled model of precedence is not rigid, however, because it is conceptually interlinked with and supplemented by a second model of leadership based on open achievement of renown and name-building through prowess in gardening and ceremonial exchange, wisdom in decision-making and dispute-settlement, and skill in establishing alliances and political support. On these skeletal criteria there is a considerable number of chiefs in Trobriand society at any given point in time. Probably the range 10-25 would quantitatively cover the number of such positio-holders at any one moment over the last eighty years.

@oc06-034 [p. 236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):

>Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society. The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111). Although usually a man, women tuai rumah are not unknown. The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival. Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement. The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.

>Jape uu me, the brave men, often the tonowis themselves, assumed the leadership during the time of hostilities. They, with other braves and elders, worked out a secret plan of the war strategy, planning the initial attack, deployment of the manpower during the offensive and later on the battle lines, they authorized and often led individual raids into the enemy territory by day or night. If need be, they organized defense of their confederacy’s villages by erecting fortifications of plank fences. In the deployment of their fighting forces they manipulated individual lineages in a similar way as our generals do army divisions. They assigned them their roles in the attack and later their positions on the battle lines.

>On the [Page 118] battle lines they were true leaders, shouting orders and expecting immediate compliance. Warriors disobeying their orders were publicly reprimanded after the battle and put to shame. Those individuals who deserted to the enemy, or otherwise committed treason (usually in-marrying wives) were put to death. Unlike our generals, the jape uu me did not stay protected in the rear, but one usually saw them in the front line and in the lead of an attacking force. By their brave behavior they acted as examples for the rest of the fighting force.

# References